\documentclass[12pt,titlepage]{article}
\usepackage{wrapfig}
\usepackage{svg}
\usepackage{nomencl}
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{geometry}
\geometry{a4paper}
\usepackage[parfill]{parskip}
\usepackage{amssymb}

\usepackage{color}
\usepackage[tt]{titlepic}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{bm}

\usepackage{listings}
\usepackage{tcolorbox}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
 
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
 
\lstset{style=mystyle}

\usepackage[section]{placeins}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{float}
\usepackage{setspace}
\usepackage[justification=raggedright]{caption}
\usepackage{sidecap}

\usepackage{xcolor}
\usepackage{framed}
\definecolor{shadecolor}{RGB}{240,240,240}


\usepackage{booktabs}
\usepackage{multirow}

\usepackage[squaren, Gray, cdot]{SIunits}
\graphicspath{{image/}} %chemin par défaut pour aller chercher les images 
\usepackage{url}
\usepackage[utf8]{inputenc}
% FOR DEFINITIONS (https://www.sharelatex.com/learn/Theorems_and_proofs)
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{definition}{Definition}[section]
\newtheorem*{definition*}{Definition}
\newtheorem{prop}{Proposition}
\newtheorem*{prop*}{Proposition}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{corollary}{Corollary}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem*{example}{Example}
\usepackage{afterpage}

\newcommand\blankpage{
\null
\thispagestyle{empty}
\newpage}

% Custom Defines
\usepackage[comma,numbers,sort&compress]{natbib}
\bibliographystyle{plainnat}
\usepackage[pdfstartview=FitH,
            breaklinks=true,
            bookmarksopen=true,
            bookmarksnumbered=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=black
            ]{hyperref}
\newcommand{\rmd}{\textrm{d}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\bi}[1]{{\ensuremath{\boldsymbol{#1}}}}
\definecolor{gray}{rgb}{0.5,0.5,0.5}

\topmargin=-0.45in      %
%\evensidemargin=0in     %
\oddsidemargin=+0.5in      %
\textwidth=5.5in        %
\textheight=9.2in       %
%\headsep=0.25in         %
\headheight=30.9pt
\linespread{1.3}

\begin{document}

% ========== TITLE PAGE ===================================================
\input{titlepage.tex}
%\onehalfspacing
  
\afterpage{\blankpage}
\clearpage{}
% ========== HEADER =======================================================
%\pagestyle{headings} 
\pagenumbering{arabic} \setcounter{page}{1}
\addtolength{\headheight}{\baselineskip}
\lhead{\textbf{Martino Milani}}
\chead{Master Thesis}
\renewcommand{\headrulewidth}{0.4pt}



\begin{abstract}

A fundamental problem in signal processing is to design computationally efficient algorithms to filter signals. In many applications, the signals to filter lie on a sphere. Meaningful examples of data of this kind are weather data on the Earth, or images of the sky. It is then important to design filtering algorithms that are computationally efficient and capable of exploiting the rotational symmetry of the problem. In these applications, given a continuous signal $f: \mathbb S^2 \rightarrow \mathbb R$ on a 2-sphere $\mathbb S^2 \subset  \mathbb R^3$, we can only know the vector of its sampled values $\mathbf f \in \mathbb R^N:\  (\mathbf f)_i = f(\mathbf x_i)$  in a finite set of points $\mathcal P \subset \mathbb S^2,\quad \mathcal P = \{\mathbf x_i\}_{i=0}^{n-1}$ where our sensors are located. Perraudin et al. in \cite{DeepSphere} construct a sparse graph $G$ on the vertex set $\mathcal P$ and then use a polynomial of the corresponding graph Laplacian matrix $\mathbf L  \in \mathbb R^{n\times n} $ to perform a computationally efficient - $\mathcal O (n)$ - filtering of the sampled signal $\mathbf f$. In order to study how well this algorithm respects the symmetry of the problem - i.e., it is equivariant to the rotation group $SO(3)$ - it is important to guarantee that the eigenvectors of $\mathbf L$  and the eigenvectors of the Laplace-Beltrami operator $\Delta_\mathbb S^2$ are somewhat ``close''.

We study the spectral properties of such graph Laplacian matrix in the special case of \cite{DeepSphere} where the sampling $\mathcal P$ is the so called HEALPix sampling (acronym for \textbf Hierarchical \textbf Equal \textbf Area iso\textbf Latitude \textbf {Pix}elization) and we show a way to build a graph $G'$ such that the corresponding graph Laplacian matrix $\mathbf L'$ shows better spectral properties than the one presented in \cite{DeepSphere}.

We investigate other methods of building the matrix $\mathbf L$ better suited to non uniform sampling measures. In particular, we studied the Finite Element Method approximation of the Laplace-Beltrami operator on the sphere, and how FEM filtering relates to graph filtering, showing the importance of non symmetric discrete Laplacians when it comes to non uniform sampling measures. We finish by showing how the graph Laplacian $\mathbf L'$ proposed in this work improved the performances of DeepSphere in a well known classification task using different sampling schemes of the sphere, and by comparing the different Discrete Laplacians introduced in this work in terms of equivariance error and filtering computational cost.


\end{abstract}

\afterpage{\blankpage}
\clearpage{}
\pagenumbering{roman}
\begin{center}
	\subsection*{ACKNOWLEDGEMENT}
\end{center}
I would first like to thank my thesis advisor Micha\"el Defferrard of the School of Electrical Engineering at EPF Lausanne. He was always there whenever I had problems or had any question. He let this work to be of my own, but he was always there to guide me in the right direction whenever it was needed. His enthusiasm and passion about research were contagious, and working with him was a real pleasure.

I would like to thank Prof. Pierre Vandergheynst of the Institute of Electrical Engineering at EPF Lausanne and Prof. Piercesare Secchi of the Department of Mathematics at Politecnico di Milano, who with their help contributed significantly to this work. Thanks to Ph.D. Nathana\"el Perraudin of the Swiss Data Science Center at ETH Z\"urich for having co-supervised my work. His intuition helped me a lot and with the numerous meetings we had he gave to this work a very valuable contribution. Thanks to Prof. Fabio Nobile of the Depatment of Mathematics at EPF Lausanne, who helped me in choosing the right direction to take when I needed it.

I would also like to thank Fr\'ed\'erik Gusset and Charles Gallay for all the interesting discussions we had in our office at the Signal Processing Laboratory at EPF Lausanne, and Luca Zampieri, for offering me a coffee whenever I needed it and for always motivating me with his incredible hard working attitude.

I then must express my deepest gratitude to my family. To my parents for having taught me to always work hard and for their unconditional support, and to my brother, for his love and for having always been by my side. This accomplishment would not have been possible without them. Thank you.

Finally, it is a pleasure to thank all my friends, those who are near and those who are far. 


\pagebreak
\null
\newpage
\pagebreak


\tableofcontents

\pagebreak

\pagenumbering{arabic}
\input{1.LiteratureReview.tex}

\pagebreak

\input{1.2.DiscreteLaplacians.tex}

\pagebreak

\input{2.Chapter1.tex}

\pagebreak

\input{3.Chapter2.tex}

\pagebreak

\input{4.Conclusions.tex}


\pagebreak
{\small \bibliography{references.bib}} 
\nocite{*}
\pagebreak

\input{5.Appendix.tex}


\pagebreak



\end{document}
